---
title: "landscape_tutorial"
author: "Ronny A. Hernández Mora"
date: "11/10/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(sp)
library(landscapemetrics)
library(sf)
library(tidyverse)
library(ggsci)
library(ggpubr)
```


## About this document

This is a tutorial that I followed from [Landscape ecology](https://bnu-landscape-ecology.netlify.app/post/2020-10-23-lsm/) in 
order to check how to use some of the GIS functions in R.

The data was downloaded from [this repo in github](https://github.com/Roberto-Supe/Landscape-Ecology-BNU-Data)

## Steps in tutorial

```{r}
beijing <- raster("data_landscape/data/map.tif")
plot(beijing)
```

To check raster attributes

```{r}
freq(beijing)
```

These are like the exploratory steps.

```{r}
check_landscape(beijing)
```

To calculate metrics, the package can use some functions inside the function
`calculate_lsm`. To know which things the package can calculate, we can use:

```{r}
glimpse(lsm_abbreviations_names)
```

For example, we want to calculate some of them from our beijing map:

```{r}
metric_class <- calculate_lsm(landscape = beijing,
                              what = c("lsm_c_ai",
                                       "lsm_c_frac_mn",
                                       "lsm_l_division",
                                       "lsm_l_pd"))
```

This will return us a data frame with the metrics.

```{r}
glimpse(metric_class)
```

This metrics can be obtained for the class, patch, landscape level.

```{r}
class_metrics <- calculate_lsm(list(beijing), # add multiple maps 
                               level = c("class"), # patch- landscape other options
                               type = "aggregation metric") # get all the metrics for the given level
```

The `landscapemetrics` package have some visualization tools:

```{r}
show_patches(beijing, labels = FALSE, class = 3) # shows the patches of class 3
```

```{r}
show_cores(beijing,labels = F,class = 2) #get the core areas of class 3 patches
```

```{r}
show_lsm(beijing,labels = F,class = 1,what = "lsm_p_area") 
# show the patches in colours based on the area,
# good to observe where that largest patch is located  
```


