---
title: "modis_data_exploration"
author: "Ronny A. Hernández Mora"
date: "11/10/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MODIStsp)
library(rgeoboundaries) # Not on CRAN
library(sf)
library(raster)
library(here)
library(ggplot2)
library(viridis)
library(leaflet)
```

## About this document

This document is to explore a waya to import and use data from MODIS and
visualize this in R.

It is based on this [documetation](https://rspatialdata.github.io/vegetation.html)

## Package

We need to install the `MODIStsp` R package. The steps to install this in
Ubuntu 20 are the following:

```{bash, eval = FALSE}
sudo apt install libjq-dev
sudo apt install libprotobuf-dev
sudo apt install protobuf-compiler
sudo apt install libv8-dev
```

Then we can install the package in R
```{r, eval = FALSE}
install.packages("V8")
install.packages("geojsonio")
install.packages("MODIStsp") # If problems persist, install "V8", then 
# geojsonio and then this one
```

## Downloading the data

The first step is to check on the MODIS product web page, select the product
we want to work with, and then we can check the layers, spectral indexes etc
that can be included in that product. In this case, we are selecting the
[aqua and terra products](https://modis.gsfc.nasa.gov/data/dataprod/mod13.php)
```{r}
MODIStsp_get_prodlayers("M*D13Q1")
```

We are going to generate a NDVI map of Costa Rica. For this, we need first
the boundaries for the country:
```{r}
# Downloading the country boundary of Mongolia
map_boundary <- geoboundaries("CostaRica")

# Defining filepath to save downloaded spatial file
spatial_filepath <- "data_vegetation/costarica.shp"

# Saving downloaded spatial file on to our computer
st_write(map_boundary, paste0(spatial_filepath))
```

Now that we have a shp file, we can download the MODIS data specifically for
the area (Costa Rica) that we selected. Everything is going to be save in
our `data_vegetation` folder.
```{r}
MODIStsp(gui             = FALSE,
         out_folder      = "data_vegetation",
         out_folder_mod  = "data_vegetation",
         selprod         = "Vegetation_Indexes_16Days_1Km (M*D13A2)",
         bandsel         = "NDVI", 
         user            = "mstp_test" ,
         password        = "MSTP_test_01",
         start_date      = "2020.06.01", 
         end_date        = "2020.06.01", 
         verbose         = FALSE,
         spatmeth        = "file",
         spafile         = spatial_filepath,
         out_format      = "GTiff")
```

After we hve the data, we can plot this:

```{r}
# Reading in the downloaded NDVI raster data
NDVI_raster <- raster(here::here("data_vegetation/costarica/VI_16Days_1Km_v6/NDVI/MYD13A2_NDVI_2020_153.tif"))

# Transforming the data
NDVI_raster <- projectRaster(NDVI_raster, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# Cropping the data
NDVI_raster <- raster::mask(NDVI_raster, as_Spatial(map_boundary))

# Dividing values by 10000 to have NDVI values between -1 and 1
gain(NDVI_raster) <- 0.0001

# Converting the raster object into a dataframe
NDVI_df <- as.data.frame(NDVI_raster, xy = TRUE, na.rm = TRUE)

# Visualising using ggplot2
ggplot() + 
  geom_raster(data = NDVI_df,
              aes(x = x, y = y, fill = MYD13A2_NDVI_2020_153)) +
  geom_sf(data = map_boundary, inherit.aes = FALSE, fill = NA) +
  scale_fill_viridis(name = "NDVI") +
  labs(title = "NDVI (Normalized Difference Vegetation Index) in Costa Rica",
       x = "Longitude",
       y = "Latitude") +
  theme_minimal()
```

We can make this plot interactive with the `leaflet` package:

```{r}
# Defining color palette
pal <- colorNumeric(c("#440154FF", "#238A8DFF", "#FDE725FF"),
                    values(NDVI_raster), 
                    na.color = "transparent")

# Visualising using leaflet
leaflet() %>%
  addTiles() %>%
  addRasterImage(NDVI_raster, colors = pal) %>%
  addLegend(pal = pal, values = values(NDVI_raster),
            title = "NDVI")
```

## References:

https://rspatialdata.github.io/index.html

https://modis.gsfc.nasa.gov/data/dataprod/mod13.php

https://docs.ropensci.org/MODIStsp/reference/MODIStsp.html

Runfola D, Anderson A, Baier H, Crittenden M, Dowker E, Fuhrig S, et al. (2020) geoBoundaries: A global database of political administrative boundaries. PLoS ONE 15(4): e0231866. https://doi.org/10.1371/journal.pone.0231866


