---
title: "Untitled"
author: "Ronny A. Hernández Mora"
date: "13/09/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(terra)
library(dplyr)
library(stringr) # for working with strings (pattern matching)
library(tidyr)   # for unite() and separate()
library(spData)
```


```{r}
dim(world)
nrow(world)
ncol(world)
```

To extract the attribute data from  an `sf`  object, we need to remove the
geomeetry

```{r}
world_df <- st_drop_geometry(world)
```

All the workflow for sf objects can be done with the tidyverse, but there are
son pitfalls that we should avoid. This are documented in the vignettes.

```{r}
sel_area <- world$area_km2 < 10000

small_coutries <- world[sel_area, ]
```

Todos los ejemplos y explicaciones vienen siendo introductorios de dplyr y tidyr.
Básicamente todo se puede manejar con lo que ya conocemos, tomando en cuenta
unicamente la columna que trae las geometrías.

## Manipulating raster objects

This is different than the vector data model, given than a raster data represents
continuos surfaces. Vector data is simple features that represents points, lines,
and polyggons as discrete entities in space.

This is a demonstration of how to create raste data:

```{r}
elev = rast(nrows = 6, ncols = 6, resolution = 0.5, 
            xmin = -1.5, xmax = 1.5, ymin = -1.5, ymax = 1.5,
            vals = 1:36)
```

```{r}
grain_char = sample(elev, 36, replace = TRUE)
grain_fact = factor(grain_char, levels = c("clay", "silt", "sand"))
grain = rast(nrows = 6, ncols = 6, resolution = 0.5, 
             xmin = -1.5, xmax = 1.5, ymin = -1.5, ymax = 1.5,
             vals = grain_fact)
```

In order to do subsets, we have to use the base R sintax `[`

```{r}
elev[1, 1]
```

To extract all values or complete rows:

```{r}
r_multi <- c(elev, grain)

# three ways to extract a layer of a stack
terra::subset(r_multi, "elev")
global(elev, sd)
hist(elev)
```


# Makking maps with R

```{r}
library(sf)
library(raster)
library(dplyr)
library(spData)
library(spDataLarge)
library(tmap)    # for static and interactive maps
library(leaflet) # for interactive maps
library(ggplot2) # tidyverse data visualization package
```


```{r}
tm_shape
```

